On the Functionality of the Multilayer Perceptron

### 2020-10-21 Thesis/Drafts/Buell-PHYS799-ThesisDraft0.tex
Updates to Feature Selections Section

@@

@@

\paragraph*{}The $21$ features are from each file are organized into one of two array structures. each called a \textit{design matrix}. A design matrix is an N-Dimensional array of real floating-point numbers that stores the features in a convenient way for the neural network to digest. For this hybrid neural-network, we use two matrices, $X_1$ and $X_2$, each containing data from $b$ audio file samples. each has shape given by:
\begin{equation}
\label{eqn-X1Shape}
X_1 \in \mathbb{R}^{(b \times N' \times k \times 1)}
\end{equation}
\begin{equation}
\label{eqn-X2Shape}
X_2 \in \mathbb{R}^{(b \times p)}
\end{equation}
Where $Q' \times R' \times 1$ gives the shape of a single spectrogram matrix $S$, described in sect. (\ref{subsec-spectrogram}). Similarly, $p$ is the number of features derived from the time and frequency-representation of features outlined in sec. (\ref{subsec-time}) and sec.(\ref{subsec-frequency}).The first axis, with size $b$ indicates that there are $b$ audio samples stored in that design matrix.

@@

\subsection{Spectrogram Feature}
\label{subsec-spectrogram}

@@

\paragraph*{}The features described in this section are derived from time-domain representations of each audio sample. Time space can be represented in one of three ways: 
\begin{itemize}
\item The raw waveform $s$\\
Indexed by $s[i]$ with $i \in [0,1,2,3,...,M-2,M-1]$
\item A single analysis frame $a_i$ \\
Indexed by $a_i[j]$ with $i \in [0,1,2,3,...,k-2,k-1]$ and $j \in [0,1,2,3,...,N-2,N-1]$
\item The full analysis frame matrix $A$ \\
Indexed by $A[i][j]$ with $i \in [0,1,2,3,...,k-2,k-1]$ and $j \in [0,1,2,3,...,N-2,N-1]$
\end{itemize}

@@

\paragraph*{}The time domain envelope (TDE) is a method of determining which parts of the sound wave contains most of the energy of the signal Often, we compute the RMS-Energy of each frame - a larger energy of the frame, the larger the average amplitude of the waveform in that frame. For frames with smaller amplitude, we can assume that the signal has either not begun, or is decaying.

\paragraph*{}We adapt this feature to \textbf{compute the RMS-Energy of the full waveform}. The RMS-Energy of the full waveform, $s$ is given by \cite{Olson,Virtanen}
\begin{equation}
\label{eqn-RMS}
\text{RMS}\big[ s \big] = \sqrt{\frac{1}{N} \sum_{i=0}^{N-1} s[i]^2}
\end{equation}
By using the full waveform, we acquire a rough estimate for the energy in the entirety of the audio file \cite{Liu}. For instruments with long sustain or release times, such as strings or undampened mallet percussion, we expect a comparably large waveform RMS when compared to those instruments without sustain such as plucked stings or dampened percussion

\begin{figure}[H]
\label{fig-FeatureTDE}
\begin{center}
\textcolor{red}{PLOT OF TDE FEATURES HERE}
\end{center}
\caption{Time domain envelope visualized in feature-space}
\end{figure}

@@

\paragraph*{}We adapt this feature to \textbf{compute the zero crossing rate for the full waveform}. The ZXR for the full waveform $s$ is given by \cite{Virtanen,Liu}
\begin{equation}
\label{eqn-ZXR}
\text{ZXR}\big[ s \big] = \frac{1}{2} \sum_{i=1}^{M-1} \big{|} \sign(s[i]) - \sign(s[i-1]) \big{|} 
\end{equation}
Where $\sign(x)$ returns $+1$ if $x > 0$, $-1$ if $x < 0$ and $0$ if $x = 0$. This provides a rough estimate for the average frequency in the full waveform, and can help discern clean periodic signals (low ZXR) from those that may have more noise or volatile behavior (high ZXR)\cite{Virtanen}.

\begin{figure}[H]
\label{fig-FeatureZXR}
\begin{center}
\textcolor{red}{PLOT OF ZXR FEATURES HERE}
\end{center}
\caption{Zero crossing rate visualized in feature-space}
\end{figure}

@@

\paragraph*{}The temporal center of mass (TCM) of a signal is used to compute roughly where in time the amplitude of the waveform \textit{bunches up}. We treat the full waveform array, $s$, with $M$ samples as a 1-Dimensional discrete mass distribution. The TCM of that waveform is then given:
\begin{equation}
\label{eqn-FeatureTCM}
\text{TCM}\big[ s \big] = \frac{\sum_{i=0}^{M-1}i s[i]}{\sum_{i=0}^{M-1}s[i]}
\end{equation}
This includes negative amplitude values acting as "negative masses".

\paragraph*{}The TCM turns out to be a very unique feature, and very powerful in classification due to it's variance. For instruments with heavier attacks, we expect the TCM to be a very near the start of the waveform. For instruments with long sustain times, we expect a very centrally located center of mass, and for instruments with long release times, we expect a later TCM value.

\begin{figure}[H]
\label{fig-FeatureTCM}
\begin{center}
\textcolor{red}{PLOT OF TCM FEATURES HERE}
\end{center}
\caption{Temporal center-of-mass visualized in feature-space}
\end{figure}

@@

@@

\paragraph*{}Mel Frequency Ceptstral Coefficients (MFCC's) are a very prolific feature used widly in audio classification tasks \cite{Liu,Virtanen,Zhang}. MFCC's are produced by converting a linear-frequency axis with units such as Hertz into units of Mels, which are design to account for the non-linear perception of frequency in humans. The Hertz to Mel and Mel to Hertz transforms are given \cite{Virtanen}:
\begin{equation}
\label{eqn-HztoMel}
M_f[h] = 2595 \log_{10}\big(1+ \frac{h}{700}\big)
\end{equation}
\begin{equation}
\label{eqn-MeltoHz}
H_f[m] = 700 \big(10^{(\frac{m}{2595})}-1\big)
\end{equation}
Where $M_f$ is the frequency in units of Mels, given $[h]$, a frequency in Hertz, and $H_f$ is the frequency in Hertz given $[m]$ a frequency in Mels.

\paragraph*{}Mel filters are produced by grouping frequency-space into $R$ overlapping bins called \textit{Mel Filter-banks}. In units of Mels, each filter bank, $R_i$ remains a constant width, but when shown in Hertz, the bins increase logarithmically. 
\begin{figure}[H]
\label{fig-MelFilterBanks}
\begin{center}
\textcolor{red}{Figure of a few Mel Filter Banks goes here}
\end{center}
\caption{Mel Filter Banks shown in frequency space with units of Hertz}
\end{figure}
Each of the $R$ filters is created to be $N'$ samples long, and transformed back into units of Hertz. When applied to an analysis frame in the frequency spectrum, the dot-product between the filter and the spectrum gives an approximation of the energy in that filter bank. Thus, each filter is concatenated into a matrix $M$ of shape $R \times N'$, where each row is a filter. We apply the Mel Filter banks to the spectrogram to create matrix $B$:
\begin{equation}
\label{eqn-FilterBanks}
B = S^T M^T
\end{equation}
Matrix $B$ has shape $k' \times R$.

\paragraph*{}Using $S^T$ ensure that each row of the matrix is an analysis frame in frequency space. Similarly, each column of the matrix $M$ is a appropriately scaled Mel filter-bank. This way, the matrix product in eqn. (\ref{eqn-FilterBanks}) allows that $B_{i,j}$ is the dot product between the $i$-th analysis frame and the $j$-th filter-bank. Finally, we compute the average energy across all $k'$ frames. The resulting $R$ floating-point numbers are approximation of the signal's energy in those particular frequency bands. For this project, we have chosen to use $R = 12$ filter-banks, which are all used as features.

\begin{figure}[H]
\label{fig-FeatureMFCC}
\begin{center}
\textcolor{red}{PLOT OF MFCC FEATURES HERE}
\end{center}
\caption{Mel Frequency Ceptral Coefficients in feature-space}
\end{figure}

@@

@@

\paragraph*{}Before making predictions on unlabeled data such as the Chaotic Synthesizers, we must determine that our model performs reasonably on data that it has never interacted with. The most common practice is to divide a full data set into a subset of \textit{training} samples, and \textit{testing} samples. The exact ratio of sample volume between these subsets varies depending on the task \cite{Goodfellow,Geron2,Mitchell}. We as the names imply, the training subset is used to fit the model, and we use the labeled testing data set to evaluate. Since the testing data is labeled, we can compare the classifiers predictions to the \textit{ground truth} labels.

@@

\subsection{Cross Validation}

@@

Divide Data into $K$ non-overlapping subsets $x_{i}$, each with roughly $K/N$ samples \\
$X \rightarrow \big\{ x_{0}, x_{1}, x_{2}, ..., x_{K-2},  x_{K-1} \big\}$ \\
Performance History $\leftarrow \{\}$

\FOR {$j = 0,1,2,3,...,K-2,K-1$}
	\item Reset all parameters in $F^*$ to a random "untrained" state
	\item Set aside testing data subset
	\item $X_{test} = x_{j}$
	\item Concatenate the remaining subsets into training data set
	\item $X_{train} = x_{j \neq k}$
	\item Train the model, $F^*$ with the $X_{train}$ data set
	\item Evaluate the trained model with metric function(s) $P$
	\item Store Performance $P$ in Performance History array
\ENDFOR

@@

\paragraph*{}By using cross-validation, we gain a deeper understanding of the model to ensure that the model can be properly trained consistently. This also helps counteract the possibility that the model got very lucky or unlucky with a particular set of initial conditions \cite{Geron,James}.

@@

>>>>

\textcolor{blue}{I need to clean this sub-subsection up as well. The optimizer needs a better introduction and justification}

@@

\textcolor{red}{I need to include a bit more detail \& citations on the "hybrid / multimodal" nature of this need works. Updates to follow}

@@

\paragraph*{}The spectrogram and $20$ features from each sample are organized into one of two array structures, each called a \textit{design matrix}. A design matrix is an array of real floating-point numbers that stores the features in a convenient way for the neural network to digest. For this hybrid neural-network, we use two matrices, $X_1$ and $X_2$, each containing data from $b$ audio file samples. each has shape given by:
\begin{equation}
\label{eqn-X1Shape}
X_1 \in \mathbb{R}^{(b \times N' \times k \times 1)}
\end{equation}
\begin{equation}
\label{eqn-X2Shape}
X_2 \in \mathbb{R}^{(b \times p)}
\end{equation}
Where $N' \times k \times 1$ gives the shape of a single spectrogram matrix $S$, described in sect. (\ref{subsec-spectrogram}). Similarly, $p$ is the number of features derived from the time and frequency-representation of features outlined in sec. (\ref{subsec-time}) and sec.(\ref{subsec-frequency}).The first axis, with size $b$ indicates that there are $b$ audio samples stored in that design matrix.

@@

\subsection{Spectrogram Features}
\label{subsec-spectrogram}

@@

\paragraph*{}The features described in this section are derived from time-domain representations of each audio sample.For consistency between smaples, each waveform is padded or truncated to contain exactly $M$ samples. Time space is indexed by $s[i]$ with $i \in [0,1,2,3,...,M-2,M-1]$

@@

\paragraph*{}The time domain envelope (TDE) is a method of determining which parts of the sound wave contains most of the energy of the signal Often, we compute the RMS-Energy of each frame - a larger energy of the frame, the larger the average amplitude of the waveform in that frame. For frames with smaller amplitude, it is likely that the signal has either not begun, or is decaying.

\paragraph*{}We adapt this feature to \textbf{compute the RMS-Energy of the full waveform}. The RMS-Energy of the full waveform, $s$ is given by \cite{Olsen,Virtanen}
\begin{equation}
\label{eqn-RMS}
\text{RMS}\big[ s \big] = \sqrt{\frac{1}{M} \sum_{i=0}^{M-1} s[i]^2}
\end{equation}
By using the full waveform, we acquire a rough estimate for the energy in the entirety of the audio file \cite{Liu}. For instruments with long sustain or release times, such as strings or undampened mallet percussion, we expect a comparably large waveform RMS when compared to those instruments without sustain such as plucked stings or dampened percussion  \textcolor{red}{Citation?}

\begin{figure}[H]
\begin{center}
\textcolor{red}{PLOT OF TDE FEATURES HERE}
\end{center}
\caption{Time domain envelope visualized in feature-space}
\label{fig-FeatureTDE}
\end{figure}

@@

\paragraph*{}We adapt this feature to \textbf{compute the zero crossing rate for the full waveform}. The ZXR for the full waveform $s$ is given by \cite{Virtanen,Liu}
\begin{equation}
\label{eqn-ZXR}
\text{ZXR}\big[ s \big] = \frac{1}{2} \sum_{i=1}^{M-1} \Big{|} \sign(s[i]) - \sign(s[i-1]) \Big{|} 
\end{equation}
Where $\sign(x)$ returns $+1$ if $x > 0$, $-1$ if $x < 0$ and $0$ if $x = 0$. This provides a rough estimate for the average frequency in the full waveform, and can help discern clean periodic signals (low ZXR) from those that may have more noise or volatile behavior (high ZXR)\cite{Virtanen}.

\begin{figure}[H]
\begin{center}
\textcolor{red}{PLOT OF ZXR FEATURES HERE}
\end{center}
\caption{Zero crossing rate visualized in feature-space}
\label{fig-FeatureZXR}
\end{figure}

@@

\paragraph*{}The temporal center of mass (TCM) of a signal is used to compute roughly where in time the amplitude of the waveform \textit{bunches up}. We treat the full waveform array, $s$, with $M$ samples as a discrete 1-Dimensional mass distribution. The TCM of that waveform is then given:
\begin{equation}
\label{eqn-FeatureTCM}
\text{TCM}\big[ s \big] = \frac{\sum_{i=0}^{M-1}i s[i]}{\sum_{i=0}^{M-1}s[i]}
\end{equation}
This includes negative amplitude values acting as "negative masses". \textcolor{blue}{Should I use the absolute value of the signal instead?}

\paragraph*{}The TCM turns out to be a very unique feature, and very powerful in classification due to it's variance textcolor{red}{Citation needed}. For instruments with heavier attacks, we expect the TCM to be a very near the start of the waveform. For instruments with long sustain times, we expect a very centrally located center of mass, and for instruments with long release times, we expect a later TCM value.

\begin{figure}[H]
\begin{center}
\textcolor{red}{PLOT OF TCM FEATURES HERE}
\end{center}
\caption{Temporal center-of-mass visualized in feature-space}
\label{fig-FeatureTCM}
\end{figure}

@@

\paragraph*{}An ACC can be thought of as convolving a signal with a time-delayed version of itself. This measures similarity between samples related to the time difference between them \textcolor{red}{Citation needed}. \textcolor{blue}{I may need to try a different set of ACC's for this}

@@

\paragraph*{}Mel Frequency Ceptstral Coefficients (MFCC's) are a very prolific feature used widely in audio classification tasks \cite{Liu,Virtanen,Zhang}. MFCC's are produced by converting a linear-frequency axis with units such as Hertz into units of Mels, which are design to account for the non-linear perception of frequency in humans. The Hertz to Mel and Mel to Hertz transforms are given \cite{Virtanen}:
\begin{equation}
\label{eqn-HztoMel}
M_f[h] = 2595 \log_{10}\big(1+ \frac{h}{700}\big)
\end{equation}
\begin{equation}
\label{eqn-MeltoHz}
H_f[m] = 700 \big(10^{(\frac{m}{2595})}-1\big)
\end{equation}
Where $M_f$ is the frequency in units of Mels, given $[h]$, a frequency in Hertz, and $H_f$ is the frequency in Hertz given $[m]$ a frequency in Mels.

\paragraph*{}Mel filters are produced by grouping frequency-space into $R$ overlapping bins called \textit{Mel Filter-banks}. In units of Mels, each filter bank, $R_i$ remains a constant width, but when shown in Hertz, the bins increase logarithmically. 
\begin{figure}[H]
\label{fig-MelFilterBanks}
\begin{center}
\textcolor{red}{Figure of a few Mel Filter Banks goes here}
\end{center}
\caption{Mel Filter Banks shown in frequency space with units of Hertz}
\end{figure}
Each of the $R$ filters is created to be $N'$ samples long, to match the width of the cropped spectrogram, and transformed back into units of Hertz. When applied to an analysis frame in the frequency spectrum, the dot-product between the filter and the spectrum gives an approximation of the energy in that filter bank. Thus, each filter is concatenated into a matrix $M$ of shape $R \times N'$, where each row is a filter. We apply the Mel Filter banks to the spectrogram to create matrix $B$:
\begin{equation}
\label{eqn-FilterBanks}
B = S^T M^T
\end{equation}
Matrix $B$ has shape $k \times R$.

\paragraph*{}Using $S^T$ ensures that each row of the matrix is an analysis frame in frequency space. Similarly, each column of the matrix $M$ is a appropriately scaled Mel filter-bank. This way, the matrix product in eqn. (\ref{eqn-FilterBanks}) allows that $B_{i,j}$ is the dot product between the $i$-th analysis frame and the $j$-th filter-bank. Finally, we compute the average energy across all $k$ frames. The resulting $R$ floating-point numbers are approximation of the signal's energy in those particular frequency bands. For this project, we have chosen to use $R = 12$ filter-banks, which are all used as features.

\begin{figure}[H]
\begin{center}
\textcolor{red}{PLOT OF MFCC FEATURES HERE}
\end{center}
\caption{Mel Frequency Ceptral Coefficients in feature-space}
\label{fig-FeatureMFCC}
\end{figure}

@@

\subsection{Assembling the Design Matrices}
\label{subsec-AssembleXs}

\textcolor{blue}{I would like to use some figures here to graphically represent how the feature vectors and spectrograms are concatenated into the design matrices. I may help with the geometry of the situation}

@@

\paragraph*{}Before making predictions on unlabeled data such as the Chaotic Synthesizers, we must determine that our model performs reasonably on data that it has never interacted with. The most common practice is to divide a full data set into a subset of \textit{training} samples, and \textit{testing} samples. As the names imply, the training subset is used to fit the model, and we use the labeled testing data set to evaluate. The exact ratio of sample volume between these subsets varies depending on the task \cite{Goodfellow,Geron2,Mitchell}. Since the testing data is labeled, we can compare the classifiers predictions to the \textit{ground truth} labels.

@@

\subsection{Cross Validation}
\label{subsec-XValidation}

@@

Divide Data into $K$ non-overlapping subsets $x_{i}$, each with roughly $N/K$ samples \\
$X \rightarrow \big\{ x_{0}, x_{1}, x_{2}, ..., x_{K-2},  x_{K-1} \big\}$ \\
Performance History $\leftarrow \{\}$

\FOR {$j = 0,1,2,3,...,K-2,K-1$}
	\item Reset all parameters in $F^*$ to a random "untrained" state
	\item Set aside testing data subset
	\item $X_{test} = x_{j}$
	\item Concatenate the remaining subsets into training data set
	\item $X_{train} = x_{i \neq j}$
	\item Train the model, $F^*$ with the $X_{train}$ data set
	\item Evaluate the trained model with the $X_{train}$ dataset, and compute value of metric function(s) $P$
	\item Store Performance $P$ in Performance History array
\ENDFOR

@@

\paragraph*{}By using cross-validation, we gain a deeper understanding of the model to ensure that the model can be consistently trained. This also helps counteract the possibility that the model got very lucky or unlucky with a particular set of initial conditions \cite{Geron,James}.

@@

\subsection{The Bagging Technique}
\label{subsec-Bagging}

\paragraph*{}Many learning models such as neural networks, or decision trees suffer from \textit{high variance} \cite{James}. This means that randomly splitting a full dataset into halves and fitting two a different model to each half has the potential to produce wildly different fit parameters. In the case of neural networks, high variance can be amplified by the nature of the randomization in multidimensional optimizer \cite{Geron}. To mitagate this effect, we can employ a variant of the \textit{bagging} algorithm. Bagging uses an average, or weighted average, of parameters in multiple models that have been trained on overlapping data sets.

\paragraph*{}Consider if we have $k$ models, all trained on overlapping training/testing data sets:
\begin{equation}
\text{Models} = \Big\{ F^*_{0} , F^*_{1} , F^*_{2} , ... , F^*_{k-2} , F^*_{k-1} \Big\}
\end{equation}
Each model contains a corresponding set of paramaters, $\Theta_{i}$, $i \in [0,k-1]$.
We could then aggregate the parameters according to a weighted average to create a new set of paramaters, $\Theta_{k}$ which can be used to construct a new model, $F^*_{k}$.

\paragraph*{}This algorithm is often used after $K$-Folds Cross validation, as outlined in in subsec. (\ref{subsec-XValidation}). This grants the $k$ models and corresponding parameters that could be used in the bagging technique \cite{James}. \textcolor{blue}{This is very not finished yet}
